gij bent 2

### 2025-10-01 paper/paper.qmd
More introduction rewriting

Linguistically, the effects of networks have been studied before. @milroy_linguistic_1985, who base themselves on @granovetter_strength_1973 and @rogers_communication_1971, demonstrated that in a network, both weak and strong ties are essential for the propagation of linguistic change. Strong ties (e.g. close friends and family) maintain and reinforce linguistic norms within tight-knit groups, resisting linguistic innovation. In contrast, weak ties (e.g. acquaintances) form bridges between groups, facilitate the spread of linguistic innovations. Weak ties are especially crucial for exposing individuals to linguistic features or patterns that are not present within their immediate, tightly connected group, enabling the flow of new linguistic forms across disparate groups and broadening the reach of linguistic innovations. Once exposed to these new forms, individuals connected through weak ties may adopt them and, through their strong ties, introduce them to their close-knit circles. This dynamic does not affect all individuals equally: some are positioned to introduce innovations into the network, while others play a crucial role in amplifying and spreading them.

Within the propagation of innovations, different roles can be distinguished. Innovators, who are less central and only loosely connected to others through weak ties, acquire novel ideas from outside their immediate community [@rogers_diffusion_2003: 282-283]. Early adopters, by contrast, are more centrally positioned and enjoy a higher degree of connectivity within the population. Through their many strong ties, they amplify and legitimize the innovation, spreading it efficiently across their dense local networks [@rogers_diffusion_2003: 282-283]. In this way, innovators and early adopters together bridge the gap between the periphery and the core of a community, ensuring that new linguistic forms do not remain isolated but instead gain broader social traction [@granovetter_strength_1973;@milroy_linguistic_1985]. The idea that innovation originates in the periphery of a network before it can spread has been attested before in network studies [@van_de_leemput_transformation_2024], but not whithin linguistics.

@@

To answer the two research questions stated above, we follow up on a study by Sevenants [@sevenants_zijt_2024] about *gij bent*, a linguistic innovation concerning the second person of 'to be' in Colloquial Belgian Dutch (CBD). To compute the individual variation of different language users, we will use elastic net regression, a technique suited to determining individual variation. First, we will introduce the *gij bent* CBD linguistic innovation. Then, we will introduce the dataset and describe how we modelled the linguistic profiles of the individual users with regard to this innovation using elastic net, and why we need this specific technique to answer the questions we have in this study. Next, we will show the results of our analyses, and discuss what consequences they have for usage-based alternation research in general.

>>>>

Linguistically, the effects of such networks have been studied before. @milroy_linguistic_1985, who base themselves on @granovetter_strength_1973 and @rogers_communication_1971, demonstrated that in a network, both weak and strong ties are essential for the propagation of linguistic change. Strong ties (e.g. close friends and family) maintain and reinforce linguistic norms within tight-knit groups, resisting linguistic innovation. In contrast, weak ties (e.g. acquaintances) form bridges between groups, facilitate the spread of linguistic innovations. Weak ties are especially crucial for exposing individuals to linguistic features or patterns that are not present within their immediate, tightly connected group, enabling the flow of new linguistic forms across disparate groups and broadening the reach of linguistic innovations. Once exposed to these new forms, individuals connected through weak ties may adopt them and, through their strong ties, introduce them to their close-knit circles. This dynamic does not affect all individuals equally: some are positioned to introduce innovations into the network, while others play a crucial role in amplifying and spreading them.

Within the propagation of innovations, different roles can be distinguished. Innovators, who are less central and only loosely connected to others through weak ties, acquire novel ideas from outside their immediate community [@rogers_diffusion_2003: 282-283]. Early adopters, by contrast, are more centrally positioned and enjoy a higher degree of connectivity within the population. Through their many strong ties, they amplify and legitimize the innovation, spreading it efficiently across their dense local networks [@rogers_diffusion_2003: 282-283]. In this way, innovators and early adopters together bridge the gap between the periphery and the core of a community, ensuring that new linguistic forms do not remain isolated but instead gain broader social traction [@granovetter_strength_1973;@milroy_linguistic_1985]. The idea that innovation originates in the periphery of a network before it can spread has been attested before in network studies [@van_de_leemput_transformation_2024], but not within linguistics.

@@

To answer the two research questions stated above, we follow up on a study by Sevenants [@sevenants_zijt_2024] about *gij bent*, a linguistic innovation concerning the second person of 'to be' in Colloquial Belgian Dutch (CBD). To compute the individual variation of different language users, we will use elastic net regression, a technique suited to determining individual variation. First, we will introduce the *gij bent* CBD linguistic innovation. Then, we will introduce the dataset and describe how we modelled the linguistic profiles of the individual users with regard to this innovation using elastic net, and why we need this specific technique to answer the questions we have in this study. Next, we will show the results of our analyses, and discuss what implications they have for usage-based alternation research in general.
